\chapter*{Preface}
\section*{Foreword}
This is submitted as my master's thesis in the Master's Degree programme in Informatics -- Program Development, at the University of Bergen and Bergen University College.

Through the work on this thesis I have had the opportunity to learn something completely new. The first years as an IT student introduced me to software modelling with UML; class, activity, sequence, and use case diagrams presented as a single language. It was confusing and hard to see the real use for these diagrams beyond serving as documentation. 

The introduction of model-driven engineering introduced me to software modelling in a way that one easily can understand the practical applications. Creating languages fitted for specific domains was an idea that really sparked an interest in me. To me, it is a whole new way of thinking programming. 

The DPF project has introduced me to a lot of skilled and interesting people who really cares about what they do. Working with people like this makes it a lot more fun and interesting for myself, as one always have someone to discuss a problem with.

\section*{Acknowledgements}
This thesis has been conducted by myself, but it would not be possible without the help of several people. The feedback and support from my supervisor Yngve Lamo has been invaluable. A special thanks goes to Florian Mantz for providing feedback and help on the technical side of things. I would also like to thank Adrian Rutle who (together with Florian) helped me figuring out the direction of my project. Øyvind Bech has also deserved a big thank you for always being helpful and providing me with \LaTeX{} sources for this thesis. I would like to thank the rest of the DPF project: Suneetha Sekhar, Xiaoliang Wang, and Alessandro Rossini.

Lastly, I want to thank my parents and my good friend and cousin, Håkon Botnen, who has helped me proofreading this thesis.

\begin{flushright}Bergen, 23 March 2012\end{flushright}